---
title: "Problem Set #2"
subtitle: "EH6105 - Quantitative Methods"
author: Steven V. Miller
output: pdf_document

fontfamily: cochineal
fontsize: 11pt
header-includes:
   - \linespread{1.05}
urlcolor: blue
---

This homework makes use of data available in `{stevedata}` and implies the use of `{tidyverse}` to answer the questions. `{tidyverse}` is not necessary to answer these questions though it will assuredly make the process easier. Load these two libraries to get started answering these questions.

```{r, message = F}
library(tidyverse)
library(stevedata)
```


# Modeling Coups in Africa, 1960 to 1975 (1982)

This homework will refer to the `african_coups` data set that is available in `{stevedata}`. This data set is an approximate reconstruction of Jackman's (1978) analysis on understanding the correlates of coup d'état susceptibility in Sub-Saharan African states. It additionally incorporates information from Johnson et al. (1984) about understanding the involvement of the military in successful or attempted coups along with new cases they consider in their expanded temporal domain. You should consult Jackman (1978) and Johnson et al. (1984) about these data, though you might find Jackman et al. (1986) informative/interesting about how big of a debate/controversy this was.

<!-- This homework will refer to the `nesarc_drinkspd` data set that is available in `{stevedata}`. This data set is loosely modified from Wave 1 (2001-2) of the National Epidemiologic Survey on Alcohol and Related Conditions (NESARC) data set, which, in part, asks a nationally representative survey of 43,093 Americans to say how often they consume alcohol on a usual day in which they drink alcohol. You can find out more information about the data by visiting [this part of the package's website](http://svmiller.com/stevedata/reference/nesarc_drinkspd.html), or with the following command.[^anotherhint] -->

- Jackman, Robert W. 1978. "The Predictability of Coups d'etat: A Model with African Data." *American Political Science Review* 72(4): 1262-75.
- Johnson, Thomas H., Robert O. Slater, and Pat McGowan. 1984. "Explaining African Military Coups d'Etat, 1960-1982." *American Political Science Review* 78(3): 622-40.
- Jackman, Robert W., Rosemary H. T. O'Kane, Thomas H. Johnson, Pat McGowan, and Robert O. Slater. 1986. "Explaining African Coups d'Etat." *American Political Science Review* 80(1): 225-49.

You can find out more information about the data by visiting [this part of the package's website](http://svmiller.com/stevedata/reference/african_coups.html), or with the following command.[^anotherhint] 

```r
?african_coups
```


Here's a little preview of these data.

```{r}
african_coups
```

Answer these questions. A successful answer of these question must include the R code you used to help you answer the question.

1. What is the median and mean of the coup index that Jackman (1978) calculates as the dependent variable in his statistical analysis?
2. What is the median and mean of the total military involvement score that Johnson et al. (1984) calculate as the dependent variable in their reanalysis of Jackman (1978)?
3. Get a histogram and density plot of Jackman's (1978) coup index or Johnson et al. (1984) military involvement score. Describe the distribution of this variable in two complete sentences.
4. What country has the highest coup index score calculated by Jackman (1978)? What country has the highest military involvement score calculated by Johnson et al. (1984)?
5. Jackman's (1978) analysis converts the variable for the percentage size of the largest ethnic group into a dummy variable (labeled as `c` in his model) if it is above a certain value. Show me the code you would use to create such a variable in these data.[^yesiknow]
6. Jackman (1978) likewise dichotomizes the turnout variable into a dummy variable (labeled as `p` in his model) if it is above a certain value. Show me the code you would use to create such a variable in these data.[^seeabove]
7. Jackman's (1978) primary explanatory variable of interest (labeled as `m` in his model) is one he creates, which he uses as a proxy for a phenomenon he calls "social mobilization" (by way of Deutsch [1961]). The data I provide can allow you to basically recreate this from his description on p. 1265. Show me how you'd create this variable.[^checkmycomment]
8. For this variable you created above, 1) calculate the mean and median, 2) identify the county that scores highest, and 3) identify the country that scores lowest.

<!-- 2. What is the mode, median, and mean of the `ethrace2a` variable, and why is this a ridiculous question to ask? -->
<!-- 3. What is the mode, median, and mean of the `sex` variable? What does the mean tell you for a binary variable like this? -->
<!-- <!-- 4. Recode the `educ` variable into a `collegeed` variable that equals 1 if the respondent says they have a four-year college degree or more. --> 
<!-- 5. Recode the `ethrace2a` variable into a dummy `white` variable that equals 1 if the respondent is white (not Hispanic) and 0 if the respondent is some other race or is Hispanic. -->
<!-- 6. Look at the `marital` variable. How many people in the data say they living with someone as married (i.e. "sambos", as you might call them in Sweden)? -->
<!-- <!-- 7. Get a histogram of the `age` variable and tell me what you see. --> 
<!-- 6. Get a histogram of the `s2aq8b` variable and tell me what you see. -->
<!-- 7. The `s2aq8b` variable is taken "as is" from the original data set, and communicates the number of drinks of any alcoholic type that the respondent says they consume on a given day drinking alcohol in the past 12 months. There are two problematic possible values in here that no longer make this a count variable, as is. What are they, and why are they problematic? -->
<!-- 8.  Recode this variable into a more informative `drinks_consumed` variable, given what you know about the variable from the codebook. Explain what you're doing. -->


[^anotherhint]: You really should always look at the codebook, and definitely here.
[^yesiknow]: A successful answer here will require you to consult either Jackman (1978) or Johnson et al. (1984). The particular value in question is made clear in both of these articles.
[^seeabove]: See the footnote for the previous prompt. The same issue applies.
[^checkmycomment]: Read the codebook I provide with the data, either in RStudio or on the package's website. This is a case where Jackman (1978, 1265) says one thing ("nonagricultural") but seems to mean another. Of the two literacy variables, use the one provided by Morrison et al. (1989) in their *Black Africa* handbook. The codebook basically says that anyway.